---
title: "Data Science Problems"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

## 3.  Data Wrangling

Data set of people is based names on the [1990 Census](https://catalog.data.gov/dataset/names-from-census-1990) and state population by age estimated from Census [factfinder](https://factfinder.census.gov).

```{r names_data}
load('sample_people.Rda')
```

### 3.1 Update the data frame (sample_people) 

Separate the person's name into a first and last name column.

*  Try with strsplit() and lapply() functions to select the appropriate element.

```{r}

split <- strsplit(sample_people$Name, ', ', fixed = TRUE)

lst <- unlist(split, use.names = FALSE)
m <- matrix(lst, nrow = 5000, byrow = T)
colnames(m) <- c('First_Name', 'Last_Name')
first <- m[, c('First_Name')]
last <- m[, c('Last_Name')]
df <- sample_people[, !(names(sample_people) %in% c('Name'))]
df$First_Name <- first
df$Last_Name <- last

```

*  Now try with the tidyr package using the separate() function.

```{r}

library(tidyr)
df <- separate(sample_people, Name, c('LastName', 'FirstName'), ', ', remove = T)

```

*  Show results with the head() function

```{r}
head(df)
```


### 3.2 Create a summary table by state.  

Recommend using the dplyr package with group_by() and summarize() functions.

* What is the sample average and median population by state?  Give both the frequency and relative frequency of the stats using inline code on a rmarkdown table:  

```{r}

library(dplyr)
g <- group_by(sample_people, State)
c <- as.data.frame(table(g$State, dnn = "State"), responseName = "Pop")
stat1 <- mean(c$Pop)
stat2 <- median(c$Pop)

r_freq <- function(x){x/5000}
c$R_Freq <- sapply(c$Pop, r_freq)
stat3 <- mean(c$R_Freq)
stat4 <- median(c$R_Freq)

```

Statistic  | Average | Median
-------------  | ------------- | -------------
Frequency  | **`r stat1`** | **`r stat2`**
Relative Frequency  | **`r stat3`** | **`r stat4`**

* Order the summary table from largest to smallest population.

```{r}

ordered <- c[order(c$Pop, decreasing = T),]

```

* [Add a html table](https://rmarkdown.rstudio.com/lesson-7.html) of the 5 larges states using kable()

```{r, results='asis'}

library(knitr)
kable(ordered[1:5,], caption = "Top 5 Populations")

```

* Add a plot of all the state populations using plot() 

```{r}

plot(ordered$State, ordered$Pop, "h")

```

* Describe the sample population data based on the table and the plot including the probability of an individual in the sample being from California.

**The sample population data shows that there was a higher number of individuals from California more than any other state. The propbability of an individual being from California is approximately 12.94%. Since California is the most populus state, this is to be expected based from our sample.**

### 3.3 Add the R default data state.name & state.region

*  Create reference data frame with state.name & state.region
*  Add the region to the sample_people data frame with the merge() function.
*  Sample one person, what is his/her percentile based on their age?
*  For that person, what is his/her percentile based on their age for his/her region?
*  For that person, what is his/her percentile based on their age for his/her state?
*  Using the mutate(), create columns for each persons age percentile for US, their region and their state.  Validate calculation based on the person sampled.
