---
title: "Data Science Problems"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

## 3.  Data Wrangling

Data set of people is based names on the [1990 Census](https://catalog.data.gov/dataset/names-from-census-1990) and state population by age estimated from Census [factfinder](https://factfinder.census.gov).

```{r names_data}
load('sample_people.Rda')
```

### 3.1 Update the data frame (sample_people) 

Separate the person's name into a first and last name column.

*  Try with strsplit() and lapply() functions to select the appropriate element.
*  Now try with the tidyr package using the seperate() function.
*  Show results with the head() function 

### 3.2 Create a summary table by state.  

Recommend using the dplyr package with group_by() and summarize() functions.

* What is the sample average and median population by state?  Give both the frequency and relative frequency of the stats using inline code on a rmarkdown table:  

Statistic  | Average | Median
-------------  | ------------- | -------------
Frequency  | _STAT1_ | _STAT2_
Relative Frequency  | _STAT3_ | _STAT4_

* Order the summary table from largest to smallest population.  
* [Add a html table](https://rmarkdown.rstudio.com/lesson-7.html) of the 5 larges states using kable() 
* Add a plot of all the state populations using plot() 
* Describe the sample population data based on the table and the plot including the probability of an individual in the sample being from California.

### 3.3 Add the R default data state.name & state.region

*  Create reference data frame with state.name & state.region
*  Add the region to the sample_people data frame with the merge() function.
*  Sample one person, what is his/her percentile based on their age?
*  For that person, what is his/her percentile based on their age for his/her region?
*  For that person, what is his/her percentile based on their age for his/her state?
*  Using the mutate(), create columns for each persons age percentile for US, their region and their state.  Validate calculation based on the person sampled.
