---
title: "Data Science Problems"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.  Data Manipulation

College data sourced from the goverment [College Scorecard](https://collegescorecard.ed.gov/) using [data API](https://collegescorecard.ed.gov/data/documentation/https://collegescorecard.ed.gov/data/) using the rscorecard package.

```{r scorecard_data}
load("college_scorecard_ca.Rda")
```

### 2.1 Create a column called npt that is either public or private average price depending on data available

*  How many CA college prices are not available?
*  What is the average price for California colleges?
*  What is the standard deviation for price?

### 2.2 Filter the data for Colleges without NAs for average price and median earnings

* What are the summary() statistics for median college salary? 
* What is the median salary for William Jessup? 
* What is the standardized score (z score) for William Jessup median earnings? 
* Using a normal distribution and the z score for William Jessup, What is the median earnings percentile for William Jessup? 
* Using the data to calculate the number of colleges less than or equal to William Jessup divided the total number of CA colleges, what is the percentile for William Jessup? 
* What is the correlation (cor) between average price and median earings? 

### 2.3 Research the US Department of Education [college scorecard](https://collegescorecard.ed.gov/)

*  Validate a few data colleges cost and median earning
*  What do you expect are other variables that impact future earnings?
