---
title: "Data Science Problems"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 2.  Data Manipulation

College data sourced from the goverment [College Scorecard](https://collegescorecard.ed.gov/) using [data API](https://collegescorecard.ed.gov/data/documentation/https://collegescorecard.ed.gov/data/) using the rscorecard package.

```{r scorecard_data}
load("college_scorecard_ca.Rda")
```

### 2.1 Create a column called npt that is either public or private average price depending on data available

```{r}
df <- college_scorecard_ca_latest
df$npt <- df$npt4_pub
na_prices_count <- sum(is.na(df$npt))
avg_price <- mean(df$npt[!is.na(df$npt)])
sd_price <- sd(df$npt[!is.na(df$npt)])
```

*  How many CA college prices are not available? **`r na_prices_count`**
*  What is the average price for California colleges? **`r avg_price`**
*  What is the standard deviation for price? **`r sd_price`**

### 2.2 Filter the data for Colleges without NAs for average price and median earnings

```{r message=FALSE}
library("dplyr")
df_filter <- df %>%
  filter(!is.na(npt)) %>%
  filter(!is.na(md_earn_wne_p10))
df_summary <- summary(df$md_earn_wne_p10)

# WJU Median Salary
wju <- df %>%
  filter(instnm == "William Jessup University")
wju_mcs <- wju$md_earn_wne_p10

# WJU Z-Score
wju_z <- (wju$md_earn_wne_p10 -   mean(df$md_earn_wne_p10[!is.na(df$md_earn_wne_p10)]))/sd(df$md_earn_wne_p10[!is.na(df$md_earn_wne_p10)])

# Percentile with pnorm
earn_filter <- df %>%
  filter(!is.na(md_earn_wne_p10))
avg_md_earn <- mean(earn_filter$md_earn_wne_p10)
sd_md_earn <- sd(earn_filter$md_earn_wne_p10)
perc <- pnorm(wju_z)

# WJU Percentile
count_under <- length(which(earn_filter$md_earn_wne_p10 <= wju$md_earn_wne_p10))
count_all <- length(earn_filter$md_earn_wne_p10)
wju_percentile <- count_under/count_all*100

# Median Earnings and Avg Cost Correlation
corr <- cor(df_filter$npt, df_filter$md_earn_wne_p10)
```

* What are the summary() statistics for median college salary? 
```{r}
df_summary
```
* What is the median salary for William Jessup? **`r wju_mcs`**
* What is the standardized score (z score) for William Jessup median earnings? **`r wju_z`**
* Using a normal distribution and the z score for William Jessup, What is the median earnings percentile for William Jessup? **`r perc*100`%**
* Using the data to calculate the number of colleges less than or equal to William Jessup divided the total number of CA colleges, what is the percentile for William Jessup? **`r wju_percentile`%**
* What is the correlation (cor) between average price and median earings? **`r corr`**

### 2.3 Research the US Department of Education [college scorecard](https://collegescorecard.ed.gov/)

*  Validate a few data colleges cost and median earning
*  What do you expect are other variables that impact future earnings?
