---
title: "4.  Data Exploration"
author: "Eric Miedema"
output:
  ioslides_presentation: default
  powerpoint_presentation:
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 2nd Project Assignment {.smaller}

Pick a city in the US:  Congratulations you are now the new mayor and want to draft a report to understand your city and set city goals.

* Organize and post city related data set.
* Create at least 4 graphs to represent city trends, demographics or other statistics.
* Select at least 2 statistics and add a goal based on a top quartile for other cities.
* Create report with city description, city statistics and goals.
* Graphs need appropriate labels and captions.

## Ideas to explore

Watch a TED talk about [cities](https://www.ted.com/topics/cities)  

*  Search "city kpis" or "city performance measures" or "city statistics" 
*  Investigate economic, social, public safety, education or health indicators

## Load city data

To get started lets use the [maps](https://cran.r-project.org/web/packages/maps/maps.pdf) package to a list of US cities.

```{r}
library(maps)
sample(us.cities$name,5)
```

## Add weather stations data

Weather data is available from [NOAA](https://www.ncdc.noaa.gov/ghcn-daily-description) and conveniently there is a rnoaa package to pull available data for [weather stations](https://www.ncdc.noaa.gov/cdo-web/datatools/findstation).

```{r}
library("rnoaa")
#station_data <- ghcnd_stations()
#station_data <- station_data[grep('^US',station_data$id),]
#save(station_data,file = 'stations_us.Rda')
load('stations_us.Rda')
sample(station_data$name,5)
```

There are `r length(unique(station_data$id))` stations so lets filter before we pull weather data.

## Filter for stations with specific elements

Lets pull stations in US states that have temperature and precipitation data in 2019.

```{r}
element_list<-c('TMAX','TMIN','PRCP','SNOW')
sd2<-station_data[station_data$element %in% element_list,]
sd2<-sd2[sd2$last_year == 2019 & sd2$state %in% state.abb,]
library(dplyr)
sd2<-sd2 %>% group_by(id) %>% 
  summarize(elements = length(element)) %>% filter(elements > 2)

us_stations<-sd2$id
```

We still have `r length(unique(us_stations))` and only `r length(unique(us.cities$name))`.

## Filter for stations close to cities

```{r}
sd3<-unique(station_data[station_data$id %in% us_stations, 
                         c('id','state','latitude','longitude','elevation')])
sd3$MERGE<-1
us_cities<-us.cities
us_cities$MERGE<-1

# Can also use dplyr::full_join
sd_cities<-merge(us_cities,sd3, by = 'MERGE', all=TRUE)
sd_cities$long_diff<-abs(sd_cities$long-sd_cities$longitude)
sd_cities$lat_diff<-abs(sd_cities$latitude-sd_cities$lat)
sd_cities$tot_diff<-sd_cities$lat_diff + sd_cities$long_diff
length(unique)
sd_cities<-sd_cities[sd_cities$tot_diff<1,]
```

Now we have `r length(unique(sd_cities$id))` and only `r length(unique(us.cities$name))`.  Can we get within 20 miles?

## Filter within 20 miles

```{r}
library(geosphere)
dist_fun<- function(x,y,a,b){
  distm(c(x, y), c(a, b), fun = distHaversine)
}
sd_cities$distance<-mapply(sd_cities$long, sd_cities$lat, 
                           sd_cities$longitude, sd_cities$latitude
                           ,FUN = dist_fun)
# convert meters to miles
sd_cities$distance_miles<-sd_cities$distance/1609.344
city_stations<-sd_cities[sd_cities$distance_miles<20,
                            c('name','id','distance_miles')]
colnames(city_stations)<-c('city','station_id','distance_miles')
#save(city_stations,file = 'city_stations.Rda')
```

Now we have `r length(unique(city_stations$station_id))` for `r length(unique(city_stations$city))`.

## Pull weather data

```{r}
station_data <- unique(city_stations$station_id)
#weather_us <- meteo_pull_monitors(station_data,
#                                      date_min = '2018-12-01',
#                                      date_max = '2018-12-31')
#save(weather_us,file = 'weather_us.Rda')
load('weather_us.Rda')
```

## Finally lets filter our city data set

```{r}
cities_weather<-unique(city_stations$city)

us_cities<-us_cities[,1:5]
colnames(us_cities)<-c('city','state','pop','lat','long')
us_cities<-us_cities[us_cities$city %in% cities_weather,]
save(us_cities,file = 'us_cities.Rda')
```

For homework 4 and project 2 this will be the starting point for cities available.