---
title: "Data Science Problems"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## 4.  Data Exploration

Congratulations!  You have been hired as an analyst for a community development company that plans and develops 55+ communities for retirees.  The first report you are assigned to work on is an exploratory analysis for a new location in the US.

```{r}
## Recommended packages
library(ggplot2)
library(gridExtra)

## weather station data
load('weather_us.Rda')
load('stations_us.Rda')
load("city_stations.Rda")

## city data
load('us_cities.Rda')

## functions
temp_convert<- function(x){ (x/10)*(9/5) + 32}
```

### 4.1  Review the US city data

*  Plot a box plot of us_cities population (pop).  What do you observe about the data?
*  Retirees are interested in living close to a city that is not to large or to small.  Exclude cities less than 1st quartile and above 3rd quartile.  Name the new data frame us_cities_mid.
*  Plot a bar chart of mid size cities by state.  What 3 states have the most mid sized cities?

### 4.2  Review weather data

* Tidy up the data for analysis
    + Filter weather_us for the first 5 columns (id, data, prcp, tmax, tmin)
    + Create a data frame of **unique** stations with the elevation, latitude and longitude using station_data. 
    + **Merge** or **Join** the station data and city station data into the weather_us data frame.
* Convert units
    +  tmax and tmin using the temp_convert function
    +  Filter extreme weather with min temp < -30 and  max temp > 120
    +  elevation from meters to feet.  Multiply by 3.281
* Create 4 scatter plots for January (low temperature) and July (high temperature).  Try geom_point() with alpha = 0.1 due to amount of data.
    +  January low temp vs latitude
    +  January low temp vs elevation
    +  July high temp vs latitude
    +  July high temp vs elevation
* Combine plots into one graphic.  Try grid.arrange(p1, p2, p3, p4, ncol=2, nrow=2).
    +  Describe the relationship between temperature, latitude and elevation.
    +  What happens after an elevation of about 1000 ft?

### 4.3 Recommendation list.

*  Summarize weather data with grouping by city
    + Average min temp from January
    + Average max temp from July
    + Average elevation
    + Add each variable to the us_cities_mid from question one.
*  Filter the data for good weather temperature.  
    + Cold temperature > 32
    + Hot temperature < 100
    + Plot a bar chart of good weather mid sized cities by state.  How does this compare to all mid sized cities in question 1.
*   Assume the targeted demographic is interested in living in a higher elevation (> 1000 ft).  What is the top recommended location?