---
title: "Project 2 - NY Unemployment"
author: "Devon Rudd"
date: "April 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Filter and Merge

The data was recieved from https://catalog.data.gov/dataset/local-area-unemployment-statistics-beginning-1976 in .csv format.

```{r}
load("us_cities.Rda")
ny_unemployment <- read.csv("Local_Area_Unemployment_Statistics__Beginning_1976.csv")
```

### Filtering us_cities

In order to match the data to make it suitable for merging, I had to first filter the data provided from moodle to only be cities from New York.

```{r}
cities <- us_cities
ny_cities <- cities[cities$state == "NY",]
```

Since each city in the provided data also had the state in the string, I removed the state abbreveation from the city string in the code below. I utilized the stringr library which makes string manipulation in R very simple.

```{r echo=FALSE}
library(dplyr)
library(stringr)

ny_cities$city <- ny_cities$city %>%
  str_remove_all(" NY")
```

### Filtering Unemployment Data

The data received for the unemployment data had Area data involved with it as well. This included counties and other state areas. I filtered the data to only keep cities and towns. I then removed the strings " City" and " Town" from each row.

```{r}
# keep only cities, eliminate county data
ny_unemployment <- ny_unemployment %>%
  filter(str_detect(ny_unemployment$Area, c("City", "Town"))) 
    # https://sebastiansauer.github.io/dplyr_filter/
  
# remove city/town from string
ny_unemployment$Area <- ny_unemployment$Area %>% 
  str_remove_all(" City") %>%
  str_remove_all(" Town")
```

### Making things match

I wanted to make sure the data had the same cities for both, so I used the cities in the provided data to filter the cities in the data retrieved.

```{r echo=FALSE}
ny_filtered <- ny_unemployment[ny_unemployment$Area == unique(ny_cities$city),]
```

### Merging

From here I was able to merge the data leaving 

```{r}
ny_all <- merge(ny_cities, ny_filtered, by.x = "city", by.y = "Area")
```

This left me with **`r print(length(unique(ny_all$city)))`** cities to individually work with. The final data head() is listed below and was saved as an Rdata file, provided on Moodle.

```{r} 
head(ny_all)
```

## Visualizing Data

Below is a histogram of the unemployed for the city of Albany, NY.

```{r}
library(ggplot2)
city = "Albany"
city_stat <- ny_all[ny_all$city == city,]
avg_ue_count <- round(mean(city_stat$Unemployed), 0)

# https://stackoverflow.com/questions/39679057/r-ggplot2-stat-count-must-not-be-used-with-a-y-aesthetic-error-in-bar-graph
ue_hist <- ggplot(data=city_stat, aes(x = Year, y = Unemployed)) +
  geom_histogram(stat = "identity", fill = 'red', color = "black") +
  geom_hline(yintercept = avg_ue_count) +
  labs(title = city)
ue_hist
```

The average number of people unemployed through the years is **`r avg_ue_count`**. Next I wanted to visualize the trend of unemployment vs. the available labor force. The graphs are shown below.

```{r}
avg_ue_rate <- round(mean(city_stat$Unemployment.Rate), 2)
ue_line <- ggplot(data=city_stat, aes(x = Year, y = Unemployment.Rate)) +
  geom_line(stat = "identity", color = "red") +
  geom_hline(yintercept = avg_ue_rate, color = 'darkblue') +
  labs(y = "Number of people", title = "Unemployment Rate")

avg_lf <- round(mean(city_stat$Labor.Force), 0)
lf_line <- ggplot(data=city_stat, aes(x = Year, y = Labor.Force)) +
  geom_line(stat = "identity", color = "blue") +
  geom_hline(yintercept = avg_lf) +
  labs(y = "Number of People", title = "Labor Force")

library(gridExtra)
grid.arrange(
  ue_line, lf_line,
  widths = c(2,1),
  layout_matrix = rbind(c(1),
                        c(2))
)
```

From this graph we can see that the unemployment rate has improved over the years, but the labor force of Albany lost a lot of employed people between the years of 1998 and 2002. The vertical lines represent the mean of each graph. For the unemployment rate, the mean was **`r avg_ue_rate`%**. The mean labor force was **`r avg_lf`**.

